---
title: "Reproducible documents"
author: "Johanne Sloth Lauszus"
format: html
---

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
```
## About me
- Johanne
- AU
- LMT

I am a *phd student* and I am **very** new to R

## Simple code

```{r}
3 * 3
```

```{r}
2 + 2
```

## Testing of git
I just want to test this 

## Showing the data
```{r}
post_meal_data
glimpse(post_meal_data)
```
## PLot one continuous variable

```{r}
#| message: false
#| warning: false
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
## Exercise: discrete plots
```{r}
#| fig-cap: "Glykemi among patients"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```
See @fig-glykemi-bar above for a cool plot!

## Plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
## side by side plots
```{r}
#| message: false
#| warning: false
#| label: fig-bmi-glycemia
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables
```{r}
#| message: false
#| warning: false
#| fig-cap: "BMI and area under the curve"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```
## Selecting data
```{r}
select(post_meal_data, Age, BMI, Weight)
```
```{r}
select(post_meal_data, -Group)
```
## Select helpers
```{r}
select(post_meal_data, starts_with("PG"))
```
```{r}
select(post_meal_data, ends_with("DXA"))
```
```{r}
select(post_meal_data, contains("0"))
```
## Renaiming columns
```{r}
small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```
## Select and rename tasks
```{r}
post_meal_data %>%
  select(contains("OGTT"))
```
```{r}
post_meal_data %>%
  select(
    contains(".."),
    -contains("...")
  )
```

```{r}
post_meal_data %>%
  rename(
    P.GLucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )
```
```{r}
post_meal_data %>%
  select(id, BMI, Length, Weight, Age) %>%
  rename(Height = Length)
```
